Approximation as Differentiation

### 2025-05-07 main/formal-structures.tex
Biproducts and semi-additive categories.

\begin{definition}[Biproducts and Semi-additive categories]
A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
and injections) give an isomorphism
\[\textstyle X \coprod Y \iso X \prod Y\] that is natural in both variables. The product/coproduct is called a
\emph{biproduct}, and the biproduct structure is denoted by $(\biprod, 0)$. The unit $0$ is both terminal and
initial and is called a \emph{zero} object.
\end{definition}

A semi-additive category $\cat{C}$ is enriched in $\CMon$, the category of commutative monoids: for any two
morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure provides a way to ``add'' them together,
forming a morphism $f + g: X \to Y$. Diagrammatically:

\begin{center}
\begin{tikzcd}
   X \arrow[r, "\diag"] & X \biprod X \arrow[r, "f \biprod g"] & Y \biprod Y \arrow[r, "\codiag"] & Y
\end{tikzcd}
\end{center}

\noindent Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the
product and $\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the
coproduct. The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \pi_1}{g \comp \pi_2} =
\coprodM{\inj_1 \comp f}{\inj_2 \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by
composing the unique maps in and out of the zero object:

>>>>

\subsection{Biproducts and Semi-Additive Categories}

\begin{definition}[Semi-additive category]
A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
and injections) give an isomorphism
\[\textstyle X \coprod Y \iso X \prod Y\] that is natural in both variables.
\end{definition}

The product/coproduct is called a \emph{biproduct}, and the biproduct structure is denoted by $(\biprod, 0)$.
The unit $0$ is both terminal and initial and is called a \emph{zero} object. A semi-additive category
$\cat{C}$ is enriched in $\CMon$, the category of commutative monoids: for any two morphisms $f, g: X \to Y$
in $\cat{C}$, the biproduct structure provides a way to ``add'' them together, forming a morphism $f + g: X
\to Y$. Diagrammatically:

\vspace{-2mm}
\begin{center}
\begin{tikzcd}
   X \arrow[r, "\diag"] & X \biprod X \arrow[r, "f \biprod g"] & Y \biprod Y \arrow[r, "\codiag"] & Y
\end{tikzcd}
\end{center}

Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the product and
$\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the coproduct.
The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \pi_1}{g \comp \pi_2} = \coprodM{\inj_1
\comp f}{\inj_2 \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by composing the unique
maps in and out of the zero object:
